---
title: "HW4"
output: 
- html_document
- github_document
---

## HPC

# Problem 1: Make sure your code is nice
Rewrite the following R functions to make them faster. It is OK (and recommended) to take a look at Stackoverflow and Google

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  ans <- rowSums(mat)
  ans
}


# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  ans <- t(apply(mat,1,cumsum))
  ans
}


# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "relative", check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "relative", check = "equivalent"
)
```

From the result we can see that original funciton and alternative function shows the same result. 

(The last argument, check = “equivalent”, is included to make sure that the functions return the same result.)



# Problem 2: Make things run faster with parallel computing

The following function allows simulating PI

```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```


In order to get accurate estimates, we can run this function multiple times, with the following code:

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```


Rewrite the previous code using parLapply() to make it run faster. Make sure you set the seed using clusterSetRNGStream():


```{r}
# YOUR CODE HERE
system.time({
  # YOUR CODE HERE
  ans <- # YOUR CODE HERE
  print(mean(ans))
  # YOUR CODE HERE
})


library(parallel)
system.time({
  cl=makePSOCKcluster(2L)
  clusterSetRNGStream(cl, 1231)
  ans <- unlist(parLapply(cl,1:4000, sim_pi, n = 10000))
  print(mean(ans))
  stopCluster(cl)
})

```


## SQL
Setup a temporary database by running the following chunk

```{r}
# install.packages(c("RSQLite", "DBI"))
library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")


# Download tables
film <- read.csv("/Users/sherryshen/Desktop/film.csv")
film_category<- read.csv("/Users/sherryshen/Desktop/film_category.csv")
category <- read.csv("/Users/sherryshen/Desktop/category.csv")


# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)

```

When you write a new chunk, remember to replace the r with sql, connection=con. Some of these questions will reqruire you to use an inner join. Read more about them here https://www.w3schools.com/sql/sql_join_inner.asp

Question 1
How many many movies is there available in each rating category.

```{sql, connection=con}
PRAGMA table_info(film)
```

```{sql, connection=con}
SELECT COUNT(*) AS n_movies, rating
FROM film
GROUP BY rating
```

Question 2
What is the average replacement cost and rental rate for each rating category.

```{sql, connection=con}
SELECT AVG(rental_rate),
       AVG(replacement_cost),
       rating
FROM film
GROUP BY rating
```


Question 3
Use table film_category together with film to find the how many films there are witth each category ID

```{sql, connection=con}
SELECT film_category.category_id,
       COUNT(*) AS n_films
FROM film
INNER JOIN film_category 
ON film.film_id = film_category.film_id
GROUP BY film_category.category_id
```



Question 4
Incorporate table category into the answer to the previous question to find the name of the most popular category.

```{sql, connection=con}
SELECT name, film_category.category_id, COUNT(film_category.film_id) AS n_film
FROM film_category
JOIN film ON film.film_id=film_category.film_id
JOIN category ON film_category.category_id=category.category_id
GROUP BY film_category.category_id
ORDER BY n_film DESC
```
Alternatively, since we know the category 15 has the highest number of film from question 3, we can just do: 

```{sql, connection=con}
SELECT name
FROM category
GROUP BY name
HAVING category_id=15
```
