---
title: "HW03"
output: 
- html_document
- github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instruction
Using the NCBI API, look for papers that show up under the term "sars-cov-2 trial vaccine." Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r counts}
library(stringr)
library(xml2)
library(httr)

website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]")
counts <- as.character(counts)
stringr::str_extract(counts, "[0-9,]+")

```

I found that there are 560 articles about "sars-cov2 trial vaccine". 

Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r ids}
query_ids <- GET(
  url = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db  = "pubmed",
    term = "sars-cov-2 trial vaccine",
    retmax = 1000
   )
)

# Extracting the content of the response of GET
ids <- content(query_ids)

# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids  "[[1]]" :return string vector
ids <- str_extract_all(ids, "<Id>[0-9]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- str_remove_all(ids, "<Id>|</Id>")
```

As we did in lab 7. Create a dataset containing the following:

Pubmed ID number,
Title of the paper,
Name of the journal where it was published,
Publication date, and
Abstract of the paper (if any).

```{r abstract}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list (
    db      = "pubmed",
    id      = paste(ids, collapse = ","),
    retmax  = 1000,
    rettype = "abstract"
  )
) 

# Turning the output into character vector
publications <- content(publications)
publications_txt <- as.character(publications)

pub_char_list <- xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)


#retrieve abstract
abstracts <- str_extract(pub_char_list, "<Abstract>(\\n|.)+</Abstract>")
#remove html text
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]]+>")
#remove new line and spaces
abstracts <- str_replace_all(abstracts, "\\s+"," ")


#retrieve journal name
journal <- str_extract(pub_char_list, "<Title>(\\n|.)+</Title>")
journal <- str_remove_all(journal, "</?[[:alnum:]]+>")
journal <- str_replace_all(journal, "\\s+"," ")

#retrieve pubdate
pubdate <- str_extract(pub_char_list, "<PubDate>(\\n|.)+</PubDate>")
pubdate <- str_remove_all(pubdate, "</?[[:alnum:]]+>")
pubdate <- str_replace_all(pubdate,"\\s+"," ")

#retrieve titles
titles <- str_extract(pub_char_list, "<ArticleTitle>(\\n|.)+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]]+>")
titles <- str_replace_all(titles, "\\s+"," ")

database <- data.frame(
  PubMedID  = ids,
  Journal   = journal,
  PubDate   = pubdate,
  Title     = titles,
  Abstract  = abstracts
)

library(kableExtra)
kable(database)

```

A new dataset has been added to the data science data repository https://github.com/USCbiostats/data-science-data/tree/master/03_pubmed. The dataset contains 3241 abstracts from articles across 5 search terms. Your job is to analyse these abstracts to find interesting insights.

Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?
Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.

We can see there are 5 search term in the dataset, which are "covid", "prostate cancer", "preeclampsia", "cystic fibrosis" and "meningitis". 
```{r}
library(tidyverse)
library(dplyr)
library(forcats)
library(ggplot2)
library(tidytext)

pubmed <- read_csv("/Users/sherryshen/Desktop/pubmed.txt")
pubmed <- pubmed %>%
  select(abstract,term)

pubmed %>%
  count(term, sort =TRUE)

```

For term "covid", we can see the most 5 common words before removing stop_words are: "the, of, covid, 19, and"; After removing stop words, we have "covid, 19, disease, patients, pandemic". 
```{r}
pubmed %>%
  filter(term %in% "covid") %>%
  unnest_tokens(word, abstract) %>%
  count(word, sort = TRUE) %>%
  top_n(5, n) 

#remove stopwords
library(stringr)
pubmed %>%
  filter(term %in% "covid") %>%
  unnest_tokens(word, abstract) %>%
  anti_join(stop_words)%>%
  count(word, sort = TRUE) %>%
  top_n(5, n) 



```

If we token abstract in to bi-grams, we have the following results: 

```{r}
#token bi-gram
pubmed %>%
  filter(term %in% "covid") %>%
  unnest_ngrams(ngram, abstract, n = 2) %>%
  count(ngram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(x = n, y = fct_reorder(ngram, n)))+
  geom_col()
```




For term "prostate cancer", we can see the most 5 common words before removing stop_words are: "the, of, and prostate, cancer"; After removing stop words, we have "prostate, cancer, patients, treatment, disease". 
```{r}
pubmed %>%
  filter(term %in% "prostate cancer") %>%
  unnest_tokens(word, abstract) %>%
  count(word, sort = TRUE) %>%
  top_n(5, n) 

pubmed %>%
  filter(term %in% "prostate cancer") %>%
  unnest_tokens(word, abstract) %>%
  anti_join(stop_words)%>%
  count(word, sort = TRUE) %>%
  top_n(5, n) 

```

If we token abstract in to bi-grams, we have the following results: 
```{r}
#token bi-gram
pubmed %>%
  filter(term %in% "prostate cancer") %>%
  unnest_ngrams(ngram, abstract, n = 2) %>%
  count(ngram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(x = n, y = fct_reorder(ngram, n)))+
  geom_col()
```

For term "preeclampsia", we can see the most 5 common words before removing stop_words are: "the, of, and, in, to"; After removing stop words, we have "pre, eclampsia, preeclampsia, women, pregnancy". 
```{r}
pubmed %>%
  filter(term %in% "preeclampsia") %>%
  unnest_tokens(word, abstract) %>%
  count(word, sort = TRUE) %>%
  top_n(5, n) 

pubmed %>%
  filter(term %in% "preeclampsia") %>%
  unnest_tokens(word, abstract) %>%
  anti_join(stop_words)%>%
  count(word, sort = TRUE) %>%
  top_n(5, n) 

```

If we token abstract in to bi-grams, we have the following results: 
```{r}
#token bi-gram
pubmed %>%
  filter(term %in% "preeclampsia") %>%
  unnest_ngrams(ngram, abstract, n = 2) %>%
  count(ngram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(x = n, y = fct_reorder(ngram, n)))+
  geom_col()
```

For term "cystic fibrosis", we can see the most 5 common words before removing stop_words are: "the, of, and, in, to"; After removing stop words, we have "fibrosis, cystic, patients, cd, disease". 
```{r}
pubmed %>%
  filter(term %in% "cystic fibrosis") %>%
  unnest_tokens(word, abstract) %>%
  count(word, sort = TRUE) %>%
  top_n(5, n) 

pubmed %>%
  filter(term %in% "cystic fibrosis") %>%
  unnest_tokens(word, abstract) %>%
  anti_join(stop_words)%>%
  count(word, sort = TRUE) %>%
  top_n(5, n) 

```

If we token abstract in to bi-grams, we have the following results: 
```{r}
#token bi-gram
pubmed %>%
  filter(term %in% "cystic fibrosis") %>%
  unnest_ngrams(ngram, abstract, n = 2) %>%
  count(ngram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(x = n, y = fct_reorder(ngram, n)))+
  geom_col()
```

For term "meningitis", we can see the most 5 common words before removing stop_words are: "the, of, and, in, a"; After removing stop words, we have "patients, meningitis, meningeal, csf, clinical". 
```{r}
pubmed %>%
  filter(term %in% "meningitis") %>%
  unnest_tokens(word, abstract) %>%
  count(word, sort = TRUE) %>%
  top_n(5, n) 

pubmed %>%
  filter(term %in% "meningitis") %>%
  unnest_tokens(word, abstract) %>%
  anti_join(stop_words)%>%
  count(word, sort = TRUE) %>%
  top_n(5, n) 

```

If we token abstract in to bi-grams, we have the following results: 
```{r}
#token bi-gram
pubmed %>%
  filter(term %in% "meningitis") %>%
  unnest_ngrams(ngram, abstract, n = 2) %>%
  count(ngram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(x = n, y = fct_reorder(ngram, n)))+
  geom_col()
```

In summary, we can see that before removing the stop words, the tokens shows nothing interesting. After we remove the stop words, the most 5 frequent tokens for each term show words like "patients" and "disease". 


Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

We can see that for term "covid", the 5 tokens with highest TF-IDF are "covid, pandemic, coronavirus, sars, cov"; for the term "prostate cancer", the 5 tokens with highest TF-IDF are "prostate, androgen, psa, prostatectomy, placenta"; for the term "preeclampsia", the 5 tokens with highest TF-IDF are "eclampsia, preeclampsia, pregnancy, maternal, gestational"; for the term "meningitis", the 5 tokens with highest TF-IDF are "meningitis, meningeal, pachymeningitis, csf, meninges"; for the term "cystic fibrosis", the 5 tokens with highest TF-IDF are "cf, fibrosis, cystic, cftr, sweat". The results are all different from what we have in question 1. 

```{r}
pubmed %>%
  unnest_tokens(word, abstract) %>%
  count(word, term) %>%
  bind_tf_idf(word, term, n) %>%
  arrange(desc(tf_idf))
```





