---
title: "HW03"
output: 
- html_document
- github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instruction
Using the NCBI API, look for papers that show up under the term "sars-cov-2 trial vaccine." Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r counts}
library(stringr)
library(xml2)
library(httr)

website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]")
counts <- as.character(counts)
stringr::str_extract(counts, "[0-9,]+")

```

I found that there are 560 articles about "sars-cov2 trial vaccine". 

Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r ids}
query_ids <- GET(
  url = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db  = "pubmed",
    term = "sars-cov-2 trial vaccine",
    retmax = 1000
   )
)

# Extracting the content of the response of GET
ids <- content(query_ids)

# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids  "[[1]]" :return string vector
ids <- str_extract_all(ids, "<Id>[0-9]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- str_remove_all(ids, "<Id>|</Id>")
```

As we did in lab 7. Create a dataset containing the following:

Pubmed ID number,
Title of the paper,
Name of the journal where it was published,
Publication date, and
Abstract of the paper (if any).

```{r abstract}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list (
    db      = "pubmed",
    id      = paste(ids, collapse = ","),
    retmax  = 1000,
    rettype = "abstract"
  )
) 

# Turning the output into character vector
publications <- content(publications)
publications_txt <- as.character(publications)

pub_char_list <- xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)


#retrieve abstract
abstracts <- str_extract(pub_char_list, "<Abstract>(\\n|.)+</Abstract>")
#remove html text
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]]+>")
#remove new line and spaces
abstracts <- str_replace_all(abstracts, "\\s+"," ")


#retrieve journal name
journal <- str_extract(pub_char_list, "<Title>(\\n|.)+</Title>")
journal <- str_remove_all(journal, "</?[[:alnum:]]+>")
journal <- str_replace_all(journal, "\\s+"," ")

#retrieve pubdate
pubdate <- str_extract(pub_char_list, "<PubDate>(\\n|.)+</PubDate>")
pubdate <- str_remove_all(pubdate, "</?[[:alnum:]]+>")
pubdate <- str_replace_all(pubdate,"\\s+"," ")

#retrieve titles
titles <- str_extract(pub_char_list, "<ArticleTitle>(\\n|.)+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]]+>")
titles <- str_replace_all(titles, "\\s+"," ")

database <- data.frame(
  PubMedID  = ids,
  Journal   = journal,
  PubDate   = pubdate,
  Title     = titles,
  Abstract  = abstracts
)

library(kableExtra)
kable(database)

```






